---
title: "Body Temperature Variations in Human Subjects"
author: "Ohm My Goodness"
date: "2023-10-04"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse) #duh
```

# Introduction

In this project, we investigate the variations in body temperature among human subjects based on their activity levels. We aim to answer the question: "Does body temperature vary in human subjects according to activity level?" This study is motivated by the need to understand how physical activity influences body temperature.

To frame our investigation, we will begin by reviewing relevant literature.

## Literature Review

We conducted a literature review to understand previous research on the effect of activity on body temperature. One relevant study by [Author et al., Year] found that...

# Methods

## Data Collection

We collected temperature data using an Arduino-based temperature sensor. Participants were monitored during two conditions: (1) at rest (sitting on the couch) and (2) after performing physical activity (jumping jacks). Axillary temperature was recorded using the sensor placed in the armpit.

# Results

## Data Combination

We combined all data files into one table for analysis. The table includes information on participant names, activity conditions, time (in milliseconds), and temperature (in degrees Celsius).

```{r, data}

# Find the data files
dat_files <- list.files(pattern = "couch|active.csv")

# Create an empty list to store data frames
dat_list <- list()

# Iterate through data files
for (file in dat_files) {
  metadata <- unlist(strsplit(file, "_"))
  name <- metadata[1]
  activity <- gsub(".csv", "", metadata[2])
  dat <- read_csv(file) %>%
    mutate(Name = name, Activity = activity)
  dat_list[[file]] <- dat
}

```

``` {r, combining data and summary}
# Combine data frames into one
combined_data <- bind_rows(dat_list)

# View the combined data
View(combined_data)

# Calculate mean temperature and standard deviation by activity and subject
summary_data <- combined_data %>%
  group_by(Activity, Name) %>%
  summarise(Mean_Temperature = mean(degC), SD_Temperature = sd(degC))

```

``` {plotting data}
# Create a plot to visualize the results
ggplot(summary_data, aes(x = Activity, y = Mean_Temperature, fill = Name)) +
  geom_bar(stat = "identity", position = "dodge") +
  geom_errorbar(aes(ymin = Mean_Temperature - SD_Temperature, ymax = Mean_Temperature + SD_Temperature),
                width = 0.2, position = position_dodge(0.9)) +
  labs(title = "Effect of Activity on Body Temperature",
       x = "Activity", y = "Mean Temperature (°C)",
       fill = "Name") +
  theme_minimal() +
  theme(legend.position = "top")

```

##Discussion

##Author Contributions

##References